Type Error Telemetry at Scale

### 2023-09-29 programming23/paper.tex
sec 4: 4.0 to 4.3 arghh

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FILL}{\textbf{FILL}}
\newcommand{\dotscale}[1]{\scalebox{0.72}{#1}}
\newcommand{\wideas}[2]{\makebox[\widthof{#2}][l]{#1}}
\newcommand{\twoline}[2]{\parbox[s]{1.4cm}{\flushleft#1\newline#2}}
\newcommand{\chkYes}{\dotscale{\CIRCLE}}
\newcommand{\chkMaybe}{\wideas{\dotscale{\Circle}}{\chkYes}}
\newcommand{\chkNo}{\wideas{}{\chkYes}}
\newcommand{\pct}[1]{\SI{#1}{\percent}}
\newcommand{\percentile}[1]{{#1}th~percentile}
\newcommand{\modefont}[1]{\texttt{#1}}
\newcommand{\mnocheck}{\modefont{nocheck}}
\newcommand{\mnonstrict}{\modefont{nonstrict}}
\newcommand{\mstrict}{\modefont{strict}}
\newcommand{\FS}{background}
\newcommand{\zerowidth}[1]{\makebox[0pt][l]{#1}}
\newcommand{\nK}[1]{#1\textsc{k}}
\newcommand{\stddev}[1]{[\nK{#1}]}
\newcommand{\gcell}[1]{\cellcolor{green!20}#1}
\newcommand{\ycell}[1]{\cellcolor{yellow!18}#1}
\newcommand{\ocell}[1]{\cellcolor{orange!29}#1}
\newcommand{\rcell}[1]{\cellcolor{red!30}$\!\!$#1$\!\!$}
\newcommand{\gbox}[1]{\colorbox{green!20}{#1}}
\newcommand{\ybox}[1]{\colorbox{yellow!18}{#1\vphantom{Ap}}}
\newcommand{\obox}[1]{\colorbox{orange!29}{#1\vphantom{Ap}}}
\newcommand{\rbox}[1]{\colorbox{red!30}{#1\vphantom{Ap}}}
\newcommand{\tefsfont}[1]{\textsc{#1}}
\newcommand{\tekey}{\tefsfont{te}}
\newcommand{\fskey}{\tefsfont{fs}}
\newcommand{\QALAN}{\FILL{} Alan:}
\newcommand{\panon}{pseudonymized}

@@

A close inspection of the data revealed a two anomalies
that we removed from further analysis:
\begin{enumerate}
  \item
    Some records have identical timestamps and session IDs.
    We keep the first such record and discard the rest.
    (A human would have to switch modules twice or enter several keystrokes
     within one millisecond to generate two such records legitimately.)

@@

\begin{table}[t]\centering
  \caption{Size of analyzed code: number of files, number of lines, and lines in edit range}
  \label{t:codebase-size}
  % ;; (list* (max* vv) (median < vv) mm (stddev/mean mm vv) (percentile* vv)))
  % ;; percentile* = 0.95 -- 0.99
  % #hash((editrange . (1156036 926 3007043/817 30975.40553032358 (0.95 8220) (0.96 9858) (0.97 13656) (0.98 18956) (0.99 34725)))
  %       (event-count . (6079 138 38689/135 582.5752201972966 (0.95 960) (0.96 1174) (0.97 1304) (0.98 1836) (0.99 3302)))
  %       (files . (54884 7678 55257996/4639 11853.650098459446 (0.95 40029) (0.96 42771) (0.97 45417) (0.98 48688) (0.99 51761)))
  %       (lines . (1089963 3115 38928947/5992 22364.75846303099 (0.95 18561) (0.96 25892) (0.97 27014) (0.98 29725) (0.99 50547)))
  %       (timespan . (1387897376 845648 591343604248/185705 15724745.722577972 (0.95 10696064) (0.96 12715468) (0.97 15596940) (0.98 21153924) (0.99 35450460))))
  %% NOTE: 3 records have +1M lines, all from one nocheck session, all with 3 files
  %% - session "30,762,216,447,324"
  %% NOTE: 133 records have +52K files, all from one nocheck session
  %% - session "330,172,870,224,081" 
  %% (2nd biggest has 41K files "31,650,473,913,529")

%% Files has HUGE numbers thanks to data assets. Forget it.
% Files  &              &
% 11,911 & \stddev{12}  &
%  7,678 &              &
% 51,761 &              &
%   & \includegraphics[width=0.2\columnwidth]{img/files-distribution.pdf}

%  \begin{tabular}{l@{}r@{~}l@{}r@{~}l}
%                 &  Lines &             &    Edit Range & \\
%    Mean [std]   &  6,497 & \stddev{22} &         3,680 & \stddev{31} \\
%    Median       &  3,115 &             &           926 & \\
%    \pct{99}     & 50,547 &             &        34,725 & \\
%    %% TODO x-axis starts at 1, not zero ... there are no zeros right??!
%      & \includegraphics[width=0.2\columnwidth]{img/lines-distribution.pdf}
%    & & \includegraphics[width=0.2\columnwidth]{img/editrange-distribution.pdf}
%
%  \end{tabular}

  \begin{tabular}[t]{lrrrl}
    ~          & Mean [std] & Median & \pct{99} Pct. & Distribution \\
    Lines      & 6,497 \stddev{22} & 3,115 & 50,547 & \includegraphics[width=0.14\columnwidth,valign=M]{img/lines-distribution.pdf} \\
    Edit Range & 3,680 \stddev{31} & 926   & 34,725 & \includegraphics[width=0.14\columnwidth,valign=M]{img/editrange-distribution.pdf}
    %% TODO x-axis starts at 1, not zero
    %% TODO much bigger font or much smaller plot!!!

  \end{tabular}

%% TODO figure --> table

\Cref{t:codebase-size} summarizes the size of projects in the dataset.
The top half of the figure reports the mean, standard deviation, median,
and \percentile{99} for the number of lines
and lines in edit range.\footnote{The data also contains the number of files in
each project, but these numbers are difficult to relate to typechecking because
projects may contain hundreds of files that define data assets.  For the
record, the median file count is 7,678 and the 99th percentile is 51,761 files.
}
The distributions zoom in zoomed-in distributions of the line and edit range
sizes.
For example, the x-axis of the first plot counts up to 1,000 lines and
the y-axis counts up to 1,000 telemetry records.

There is a huge amount of variation across projects.
The largest ones have over 50,000 lines of code
while the smallest have merely 1 line of code.
Unsurprisingly, the numbers come with large standard
deviations~([std] in~\cref{t:codebase-size}).
The median values are more reasonable, with roughly 3,000 lines of code
and 1,000 lines in edit ranges.

@@

\begin{figure}[t]\centering
  %% NOTE max time = 1 million seconds = 15 days ...  +100 are over 2 days

%  \begin{tabular}{l@{}r@{~}l@{}r@{~}l} \\
%                 & Time Span (sec)  &       & Record Count  \\
%    Mean [std]   &     3,184 & \stddev{16}  &     286 & [583] \\
%    Median       &       845 &              &     138        \\
%    \pct{99}     &    35,450 &              &   3,302        \\
%    %% TODO x-axis starts at 1, not zero ... there are no zeros right??!
%    & \includegraphics[width=0.2\columnwidth]{img/timespan-distribution.pdf}
%    & & \includegraphics[width=0.2\columnwidth]{img/event-count-distribution.pdf}
%
%  \end{tabular}

  \begin{tabular}{lrrrl} \\
    ~               & Mean [std] & Median & \pct{99} Pct. & Distribution \\
    Time Span (sec) & 3,184 \stddev{16} & 845 & 35,450 & \includegraphics[width=0.14\columnwidth,valign=M]{img/timespan-distribution.pdf} \\
    Record Count    & 286 [583]         & 138 & 3,302  & \includegraphics[width=0.14\columnwidth,valign=M]{img/event-count-distribution.pdf}
    %% TODO x-axis starts at 1
    %% TODO huge font, or smaller plot

  \end{tabular}

  \caption{Session size in seconds and in number of records.}
  \label{f:session-size}
\end{figure}

\Cref{f:session-size} outlines the size of sessions using two metrics:
real time in seconds between the first and last record, and the total number of
records.
The top of~\cref{f:session-size} reports the mean, standard deviation,
median, and \percentile{99}; the bottom shows size distributions, zoomed in
to focus on at most 1,000 units (x-axis) and 1,000 sessions
(y-axis).\todo{Again, confused.}

Similar to the data on project size, there is huge variation in session
length.
The longest sessions last several days and/or contain thousands of records,
while the shortest last a few milliseconds or include just one record.
Likewise, both plots in the figure follow a power-law distribution
(especially for record count).
A typical session runs on the order of minutes and consists of a few dozen events.
The median session time is approximately 14 minutes.
The median number of
records is 138.
% These numbers are reasonable.

@@

\begin{figure}[t]
  \begin{tabular}[t]{ll} \\
    \begin{tabular}[t]{r@{~~}r@{~}l}
      595,137 & type errors \\
      289,698 & in current module & [\pct{48.68}] \\
       30,924 & in edit range & [\pct{5.20}]
    \end{tabular}
    \begin{tabular}[t]{r@{~~}r@{~}l}
      72,235,735 & {\FS{} errors} \\
      37,027,281 & in current module & [\pct{51.26}] \\
       1,111,178 & in edit range & [\pct{1.54}]
    \end{tabular}
  \end{tabular}
  \caption{Type errors and \FS{} errors across all telemetry records.}
  \label{f:count-analysis-errors}
\end{figure}

\FILL{} need old and current in edit range

\Cref{f:count-analysis-errors} lists the number of type analysis errors
and breaks them down by location in the codebase.
In total, there are over 590,000 type errors.
%% delete? upcoming figure has the same info
% Most of these (\pct{65}) appeared in modules that opted in to typechecking with either the
% \mnonstrict{} or \mstrict{} mode; the remainder are syntax errors reported
% by \mnocheck{} mode.
The number of \FS{} errors is much larger, 72M, because this analysis
ran on every module using strict type.
(Furthermore, \FS{} runs silently, so creators have no
awareness of or incentive to
fix these errors.)
Approximately half the analysis errors occurred in the current module,
and a small fraction overlapped with the current edit range~(\pct{5} for
type errors, \pct{1} for \FS{}).

The large fraction of errors in the current module is encouraging.
It suggests that errors appear locally, as the result of edits to nearby
code (as opposed to edits in one module raising errors in another), and
that creators fix the errors before switching focus to another
module.
The low fraction of errors in the edit range suggests that errors rarely
persist through nearby edits.

@@

Creators have three analysis modes to choose from.
Furthermore, they can switch between modes at any time.
According to~\cref{s:type-analysis-modes}, however, usage is
extremely skewed toward the type-free \mnocheck{} mode.
Nearly \pct{90} of all records use the \mnocheck{} mode
and \pct{90} of all sessions use \mnocheck{} exclusively.
Among the rest, roughly \pct{10} of all records use \mnonstrict{} and only a
tiny fraction (\pct{0.4}) use \mstrict{}.
The per-session data is similar: \pct{9} use \mnonstrict{} exclusively
and \pct{0.2} use \mstrict{}.

These adoption numbers indicate that there is a communication gap;
as long as type checking is opt-in, most developers stay opted out.
Once Luau type checking is opt-out rather than opt-in, it will be interesting
to revisit these numbers.

%% mostly pure, rare mixes
Most sessions (\pct{99.82}) stick to a single analysis mode.
These sessions never change the mode in the current module and never switch
focus to a module with a different analysis mode.
Among the mixed-mode sessions, most (\pct{80}) switch to modules with
a different mode, about half contain at least one edit that upgrades
the typing strictness in the
current mode, and about half contain a downgrade.
There are 619 upgrades in total and 583 downgrades.
\FILL{} Ben check how often the downgrades follow the upgrades.

@@

A possible explanation for the low adoption of \mnonstrict{} and
\mstrict{} mode is that upgrading to these modes leads to a large
number of type errors.
Creators might get discouraged or overwhelmed by a high error count and revert to
\mnocheck{} mode.

The data does not support this explanation.
On average, the mode-upgrades in the data resulted in 3 additional
type errors (stddev 7, median 0).
The worst-case increase was 57 type errors, which is very high
but exceptional.
In another exception, upgrading modes removed three type errors
(possibly due to other edits being rolled in with the
mode change).

In the other direction, mode downgrades have only a small negative effect on the
number of errors (mean -0.3, stddev 4, median 0, max -48).
%% curious outlier: downgrade added 30 errors!
We would see a much larger effect here if creators used downgrades
as a quick way to silence the typechecker.
But, creators evidently switch modes only when the code is already
in well-typed shape.

%% \paragraph{Module Switches vs Errors}
%% FILL need \%s out of all module switches.
%% How many module switches exit a module with errors?
%% > OLD DATA
%% > \mnocheck{} 24, \mnonstrict{} 9, \mstrict{} 3.
%% > Rare?
%% > More common for forcedstrict:
%% >  \mnocheck{} 190, \mnonstrict{} 16, \mstrict{} 2.
%% >
%% > How many module switches go to a module with errors?
%% > \mnocheck{} 70, nonstrict 9, strict 4
%% > For FS: \mnocheck{} 400, \mnonstrict{} 39, \mstrict{} 6.

@@

  \item[How often do sessions turn analysis off?]
    Sessions rarely disable type analysis after opting in.
    Only 233 of the \nK{340} sessions ($<\pct{0.07}$)
    switch to a weaker mode.
    Most of these (\pct{75}) downgrade from \mstrict{} to
    \mnonstrict{} % BEN: confirm!
    rather than from \mnonstrict{} mode.
    %% FILL more insight if we match upgrades + downgrades in session?

    %% upgrade
    %% 519 total sessions
    %% 176 with up
    %% #hash((nocheck . 0) (nonstrict . 44) (strict . 132))
    %%  nocheck : 0.00\%
    %%  strict : 75.00\%
    %%  nonstrict : 25.00\%
    %%
    %% downgrade
    %% 515 total sessions
    %% 233 with down
    %% #hash((nocheck . 0) (nonstrict . 64) (strict . 169))
    %%  nocheck : 0.00\%
    %%  strict : 72.53\%
    %%  nonstrict : 27.47\%

>>>>

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FILL}{\textbf{FILL}}
\newcommand{\dotscale}[1]{\scalebox{0.72}{#1}}
\newcommand{\wideas}[2]{\makebox[\widthof{#2}][l]{#1}}
\newcommand{\twoline}[2]{\parbox[s]{1.4cm}{\flushleft#1\newline#2}}
\newcommand{\chkYes}{\dotscale{\CIRCLE}}
\newcommand{\chkMaybe}{\wideas{\dotscale{\Circle}}{\chkYes}}
\newcommand{\chkNo}{\wideas{}{\chkYes}}
\newcommand{\pct}[1]{\SI{#1}{\percent}}
\newcommand{\percentile}[1]{{#1}th~percentile}
\newcommand{\modefont}[1]{\texttt{#1}}
\newcommand{\mnocheck}{\modefont{nocheck}}
\newcommand{\mnonstrict}{\modefont{nonstrict}}
\newcommand{\mstrict}{\modefont{strict}}
\newcommand{\FS}{background}
\newcommand{\zerowidth}[1]{\makebox[0pt][l]{#1}}
\newcommand{\nK}[1]{#1\textsc{k}}
\newcommand{\stddev}[1]{\nK{#1}}
\newcommand{\gcell}[1]{\cellcolor{green!20}#1}
\newcommand{\ycell}[1]{\cellcolor{yellow!18}#1}
\newcommand{\ocell}[1]{\cellcolor{orange!29}#1}
\newcommand{\rcell}[1]{\cellcolor{red!30}$\!\!$#1$\!\!$}
\newcommand{\gbox}[1]{\colorbox{green!20}{#1}}
\newcommand{\ybox}[1]{\colorbox{yellow!18}{#1\vphantom{Ap}}}
\newcommand{\obox}[1]{\colorbox{orange!29}{#1\vphantom{Ap}}}
\newcommand{\rbox}[1]{\colorbox{red!30}{#1\vphantom{Ap}}}
\newcommand{\tefsfont}[1]{\textsc{#1}}
\newcommand{\tekey}{\tefsfont{te}}
\newcommand{\fskey}{\tefsfont{fs}}
\newcommand{\QALAN}{\FILL{} Alan:}
\newcommand{\panon}{pseudonymized}

@@

A close inspection of the data revealed two anomalies:
\begin{enumerate}
  \item
    Some records have identical timestamps and session IDs.
    We keep the first such record and discard the rest.
    (A human would have to switch modules twice or enter several keystrokes
     within one millisecond to generate two such records legitimately.)

@@

\begin{table}[t]\centering
  \caption{Size of analyzed code: number of files, number of lines, and lines in edit range}
  \label{t:codebase-size}
  % ;; (list* (max* vv) (median < vv) mm (stddev/mean mm vv) (percentile* vv)))
  % ;; percentile* = 0.95 -- 0.99
  % #hash((editrange . (1156036 926 3007043/817 30975.40553032358 (0.95 8220) (0.96 9858) (0.97 13656) (0.98 18956) (0.99 34725)))
  %       (event-count . (6079 138 38689/135 582.5752201972966 (0.95 960) (0.96 1174) (0.97 1304) (0.98 1836) (0.99 3302)))
  %       (files . (54884 7678 55257996/4639 11853.650098459446 (0.95 40029) (0.96 42771) (0.97 45417) (0.98 48688) (0.99 51761)))
  %       (lines . (1089963 3115 38928947/5992 22364.75846303099 (0.95 18561) (0.96 25892) (0.97 27014) (0.98 29725) (0.99 50547)))
  %       (timespan . (1387897376 845648 591343604248/185705 15724745.722577972 (0.95 10696064) (0.96 12715468) (0.97 15596940) (0.98 21153924) (0.99 35450460))))
  %% NOTE: 3 records have +1M lines, all from one nocheck session, all with 3 files
  %% - session "30,762,216,447,324"
  %% NOTE: 133 records have +52K files, all from one nocheck session
  %% - session "330,172,870,224,081" 
  %% (2nd biggest has 41K files "31,650,473,913,529")
%% Files has HUGE numbers thanks to data assets. Forget it.
% Files  &              &
% 11,911 & \stddev{12}  &
%  7,678 &              &
% 51,761 &              &
  \begin{tabular}[t]{lrrrrl}
    ~          & Mean  & Stddev & Median & \pct{99} Pct. & Distribution \\
    Lines      & 6,497 & \stddev{22} & 3,115 & 50,547 & \includegraphics[width=0.14\columnwidth,valign=M]{img/lines-distribution.pdf} \\
    Edit Range & 3,680 & \stddev{31} & 926   & 34,725 & \includegraphics[width=0.14\columnwidth,valign=M]{img/editrange-distribution.pdf}
    %% TODO x-axis starts at 1, not zero
    %% TODO much bigger font or much smaller plot!!!
  \end{tabular}
\end{table}

\Cref{t:codebase-size} summarizes the size of projects in the dataset.
The middle four columns report the mean, standard deviation, median,
and \percentile{99} for the number of lines
and lines in edit range.\footnote{The data also contains the number of files in
each project, but these numbers are difficult to relate to typechecking because
projects may contain hundreds of files that define data assets.  For the
record, the median file count is 7,678 and the 99th percentile is 51,761 files.
}
The plots on the right show zoomed-in distributions of the line and edit range
sizes.
For example, the x-axis of the first plot counts up to 1,000 lines (not showing
a long tail to the right) and the y-axis counts up to 1,000 telemetry records
(not showing a tall spike on the left).

There is a huge amount of variation across projects.
The largest ones have over 50,000 lines of code
while the smallest have merely 1 line of code.
Unsurprisingly, these wide-ranging numbers come with large standard
deviations.
The median values are more reasonable, with roughly 3,000 lines of code
and 1,000 lines in edit ranges.

@@

\begin{table}[t]\centering
  \caption{Session size in seconds and in number of records.}
  \label{t:session-size}
  %% NOTE max time = 1 million seconds = 15 days ...  +100 are over 2 days
  \begin{tabular}{lrrrrl} \\
    ~               & Mean & Stddev & Median & \pct{99} Pct. & Distribution \\
    Time Span (sec) & 3,184 & \stddev{16} & 845 & 35,450 & \includegraphics[width=0.14\columnwidth,valign=M]{img/timespan-distribution.pdf} \\
    Record Count    & 286 & 583         & 138 & 3,302  & \includegraphics[width=0.14\columnwidth,valign=M]{img/event-count-distribution.pdf}
    %% TODO x-axis starts at 1
    %% TODO huge font, or smaller plot
  \end{tabular}
\end{table}

\Cref{t:session-size} outlines the size of sessions using two metrics:
real time in seconds between the first and last record, and the total number of
records.
The middle columns report the mean, standard deviation,
median, and \percentile{99}.
Similar to the previous table~(\cref{t:codebase-size}), the distributions focus
on the lower-left fragment of the data, which again has a very long tail.
%% comment on long-tail record counts?
The longest sessions last several days and/or contain thousands of records,
while the shortest last a few milliseconds or include just one record.
% Likewise, both plots in the figure follow a power-law distribution
% (especially for record count).
A median session runs on the order of minutes and consists of a few dozen events.
%% The median time is 14 minutes and the median 
%% The median number of records is 138.
% These numbers are reasonable.

@@

\begin{table}[t]\centering
  \caption{Current type errors and \FS{} errors across all telemetry records.}
  \label{t:count-analysis-errors}
  \begin{tabular}[t]{ll} \\
    \begin{tabular}[t]{r@{~~}r@{~}r}
      595,137 & type errors \\
      289,698 & \hbox{}~in module & [\pct{48.68}] \\
       30,924 & \hbox{}~in edit range & [\pct{5.20}]
    \end{tabular}
    \begin{tabular}[t]{r@{~}r@{~}r}
      72,235,735 & {\FS{} errors} \\
      37,027,281 & \hbox{}~in module & [\pct{51.26}] \\
       1,111,178 & \hbox{}~in edit range & [\pct{1.54}]
    \end{tabular}
  \end{tabular}
\end{table}

\Cref{t:count-analysis-errors} lists the number of type analysis errors
and breaks them down by location in the codebase.
In total, there are over 590,000 type errors.
The number of \FS{} errors is much larger, at 72 million, because this analysis
uses rigorous checks and runs on every module no matter what type analysis
mode the module declares.
(Furthermore, \FS{} checks run silently, so creators have no
awareness of or incentive to fix these errors.)
Approximately half the analysis errors occurred in the current module,
and a small fraction overlapped with the current edit range~(\pct{5} for
type errors, \pct{1} for \FS{}).

The large fraction of errors in the current module is encouraging
for two reasons.
First, it suggests that errors appear locally, as the result of edits to nearby
code, rather than as the result of edits to far-away code that lives in another
module.
Second, it suggests that creators typically fix errors before switching
to another module.

The low fraction of errors in the edit range sends a mixed message:
\begin{itemize}
  \item
    On one hand, it suggests that edits successfully remove errors.
    This is true, however, only for type errors and not for \FS{} errors.
    There are 11,479 more type errors from the previous analysis
    that overlap with the current edit range~(\pct{27} difference),
    but 172,678 fewer previous \FS{} errors~(\pct{15} difference).
    Since creators can see the type errors but not the \FS{}
    errors before making edits, we conclude that the visibility
    makes a difference.
    \Cref{s:type-error-survival} explores edits and type errors in more detail.

  \item
    On the other hand, the low overlap between errors (whether current or previous)
    and edits says that edits rarely target highlighted code.
    There are several possible explanations:
    analysis might report several errors and the edits only target one or two of them;
    creators might ignore errors entirely and edit other code; or
    creators might choose to edit a non-highlighted location to fix an error.
    Our data cannot narrow down the answer.

@@

Creators have three analysis modes to choose from
and can switch between modes at any time.
\Cref{f:dataset-overview} shows, however, that
usage is extremely skewed toward the default \mnocheck{} mode.
Nearly \pct{90} of all records use \mnocheck{}, while \pct{10} use
\mnonstrict{} and only a tiny fraction~(\pct{0.4}) use \mstrict{} mode.
Grouping these records into sessions shows that
\pct{90} of all sessions use \mnocheck{} exclusively,
\pct{9} use \mnonstrict{} exclusively,
and \pct{0.2} use \mstrict{} exclusively.

These adoption numbers indicate that as long as type checking is opt-in, most
creators stay opted out.
In the future, Luau plans to make \mnonstrict{} the default mode.
It will be interesting to revisit mode usage at that point.

%% mostly pure, rare mixes
Most sessions (\pct{99.82}) stick to a single analysis mode.
They never change the mode in the current module and never
switch focus to a module with a different analysis mode.
Among the other, mixed-mode sessions, most of those (\pct{80}) switch to a
module with a different mode, about half contain at least one edit that
upgrades to a more strict mode, and about half contain a downgrade to a
less-strict mode.
There are 263 total upgrades (across all sessions) and 320 total downgrades.

@@

A possible explanation for the low adoption of \mnonstrict{} and
\mstrict{} mode is that upgrading to these modes leads to a large
number of type errors.
Creators might get discouraged or overwhelmed by a high error count.

The data does not support this explanation.
On average, mode-upgrades resulted in 3 additional
type errors (stddev 7, median 0).
The worst-case increase was quite high at 57 type errors, but exceptional.
In another exception, upgrading modes removed three type errors,
possibly due to other edits being rolled in with the
mode change.

In the other direction, mode downgrades have only a small negative effect on the
number of errors (mean -0.3, stddev 4, median 0, max -48).
%% curious outlier: downgrade added 30 errors!
We would see a much larger effect here if creators used downgrades
as a quick way to silence the typechecker.
But, creators evidently switch modes only when the code is already
in a well-typed state.

%% \paragraph{Module Switches vs Errors}
%% FILL need \%s out of all module switches.
%% count-modswitch in downgrade-count.rkt
%% How many module switches exit a module with errors?
%% > OLD DATA
%% > \mnocheck{} 24, \mnonstrict{} 9, \mstrict{} 3.
%% > Rare?
%% > More common for forcedstrict:
%% >  \mnocheck{} 190, \mnonstrict{} 16, \mstrict{} 2.
%% >
%% > How many module switches go to a module with errors?
%% > \mnocheck{} 70, nonstrict 9, strict 4
%% > For FS: \mnocheck{} 400, \mnonstrict{} 39, \mstrict{} 6.

@@

  \item[How often do sessions turn analysis off?]
    Sessions rarely disable type analysis after opting in.
    Only 233 of the \nK{340} sessions ($<\pct{0.07}$)
    switch to a weaker mode.
    One-fourth of the downgrades switch from \mstrict{}
    to \mnonstrict{}, and therefore keep some type analysis.
    The remainder switch from \mstrict{} to \mnocheck{} (\pct{50})
    or from \mnonstrict{} to \mnocheck{} (\pct{25}).
    %% FILL more insight if we match ups and downs together??

    %% upgrade
    %% 519 total sessions
    %% 176 with up
    %% 263 total up
    %% #hash((nocheck . 0) (nonstrict . 44) (strict . 132))
    %%  nocheck : 0.00\%
    %%  strict : 75.00\%
    %%  nonstrict : 25.00\%
    %%%% #hash(((strict nocheck) . 88)
    %%%%       (nonstrict . 0)
    %%%%       (strict . 0)
    %%%%       ((nonstrict nocheck) . 44)
    %%%%       (nocheck . 0)
    %%%%       ((strict nonstrict) . 44))
    %%%%  (strict nonstrict) : 25.00\%
    %%%%  nocheck : 0.00\%
    %%%%  (nonstrict nocheck) : 25.00\%
    %%%%  strict : 0.00\%
    %%%%  nonstrict : 0.00\%
    %%%%  (strict nocheck) : 50.00\%
    %%
    %% downgrade
    %% 515 total sessions
    %% 233 with down
    %% 320 total down
    %% #hash((nocheck . 0) (nonstrict . 64) (strict . 169))
    %%  nocheck : 0.00\%
    %%  strict : 72.53\%
    %%  nonstrict : 27.47\%
    %%%% #hash(((strict nocheck) . 116)
    %%%%       (nonstrict . 0)
    %%%%       (strict . 0)
    %%%%       ((nonstrict nocheck) . 64)
    %%%%       (nocheck . 0)
    %%%%       ((strict nonstrict) . 53))
    %%%%  (strict nonstrict) : 22.75\%
    %%%%  nocheck : 0.00\%
    %%%%  (nonstrict nocheck) : 27.47\%
    %%%%  strict : 0.00\%
    %%%%  nonstrict : 0.00\%
    %%%%  (strict nocheck) : 49.79\%
